Structured Versus Decorated Cospans

### 2021-06-21 cospans_new.tex
changed names of decorations in Section 2 to letters like s and t.

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $(F,\phi,\phi_0) \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

\[ \left(a\to m\leftarrow b,x\in F(m)\right), \qquad \left(b\to n\leftarrow c,y\in F(n)\right)\] 

@@

together with the decoration in $y \odot x \in F(m +_b n)$ specified by this functor:
\[  \one\cong\one \times \one \xrightarrow{x \times y} F(n) \times F(m) \xrightarrow{\phi_{n,m}} F(m+n) \xrightarrow{F(\psi)} F(m +_{b}n). \]

@@

\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};

@@

\node (J) at (7,0.5) {$y \in F(n)$};
\node (K) at (7,-0.5) {$y' \in F(n')$};
\node (L) at (1,-1) {$\tau_\alpha \maps F(h_1)(x) \to x'$};
\node (M) at (5,-1) {$\tau_\beta \maps F(h_2)(y) \to y'$};

@@

\node (E) at (4.5,0.5) {$y \odot x \in F(m+_b n)$};
\node (F) at (4.5,-0.5) {$y' \odot x' \in F(m' +_{b'} n')$};

@@

together with the decoration morphism $\tau_{\beta \odot \alpha} \maps F(h_1 +_g h_2)(y \odot x) \to (y' \odot x')$ specified by this natural transformation:

@@

(D) edge node [above] {$x \times y$} (E)
(D) edge node [below] {$x' \times y'$} (E')

@@

where the middle isomorphism is \cref{eq:pseudonaturality} from pseudonaturality of $\phi$ and the right-hand isomorphism comes from the pseudofunctoriality of $F$.

\begin{thm}\label{thm:decorated_cospans_2}
Let $(\A,+)$ be a cocartesian monoidal category and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ be a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a \to m \leftarrow b ,x \in F(m))$ and $(c \to n \leftarrow d, y\in F(n))$ is
\[
\begin{tikzcd}
& m+n & \\
a+c\ar[ur,"i_1+i_2"] && b+d,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad x \otimes y:=  \phi_{m,n}(x,y) \in F(m + n) 
\]
\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(x) \to x'\textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(y) \to y'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m,n}(x,y))\to \phi_{m',n'}(x',y')\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{h,h'}}{\cong}}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"x\times y"]\ar[dr,"x'\times y'"']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

We prove both these theorems using the work of Shulman \cite{Shulman2008}, who gives a general way to construct double categories from `Beck--Chevalley monoidal bifibrations'---a concept we explain in \cref{subsec:fibrations}.

@@

Note that when $\C$ is the arrow category of $\A$ and $\Phi$ maps any arrow in $\A$ to its
domain, an object of $\Fr(\Phi)$ simply amounts to a cospan in $\C$.  In this case $\Fr(\Phi)$ is the double category of cospans in $\C$.    We shall use variations on this idea to construct double categories of decorated and structured cospans, and to prove that under certain conditions these double categories are equivalent.    Here we consider decorated cospans, proving Theorem \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying Shulman's result to a particular monoidal bifibration built from a lax monoidal pseudofunctor $F \maps (A,+) \to (\Cat,\times)$.

@@

In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{subsec:bicats}, the induced monoidal structure on the Grothendieck category $\X=\inta F$ (\cref{def:GrothCat}) is given by
\begin{equation}\label{eq:explicitstructure2}
\Big(a,x\in F(a)\Big)\ot_{\X}\Big(b,y\in F(b)\Big)=\Big(a+b,\phi_{a,b}(x,y)\in F(a+b)\Big), \qquad I_{\X}=\Big(0_A,\phi_0\Big)
\end{equation}
If $F$ is a symmetric lax monoidal pseudofunctor, the induced monoidal structure in $\inta F$ is symmetric via
$$\left(\beta_{a,b},(u_{a,b})_{x,y}\right)\maps \left(a+b,\phi_{a,b}(x,y)\right)\simrightarrow\left(b+a,\phi_{b,a}(y,x)\right)$$ where $\beta$ is the canonical symmetry for $\A$ and $u$ is the natural isomorphism of \cref{eq:braidedpseudofun}.

Moreover, each fiber $\X_a=F(a)$ obtains a monoidal structure via 
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a)
\end{equation}
where $\nabla$ is the fold map, which is symmetric when $F$ is again via the components of $u_{a,a}$. Also, each reindexing functor $f_!=F(f)$ obtains a strong monoidal structure with
\begin{equation}\label{eq:strongreindexing}
\begin{tikzcd}[row sep=.35in]
 F(a)\times F(a)\ar[r,"\phi_{a,a}"]\ar[d,"F(f)\times F(f)"']\ar[dr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{f,f}}{\cong}}"] & 
 F(a+a)\ar[r,"F(\nabla)"]\ar[d,"F(f+f)"description]\ar[dr,phantom,"\cong"] & F(a)\ar[d,"Ff"] \\
 F(b)\times F(b)\ar[r,"\phi_{b,b}"'] & F(b+b)\ar[r,"F(\nabla)"'] & F(b)
\end{tikzcd}
\end{equation}
\end{proof}

Now we are ready to use Shulman's result to prove Theorems \ref{thm:decorated_cospans_1} and  \ref{thm:decorated_cospans_2}.  An $F$-decorated cospan from $a$ to $b$ 
is a cospan $a \to m \leftarrow b$ in $\A$ together with an object $x \in F(m)$.   This can be seen as a pair $(e \to m, x \in F(m))$ such that $e$ equals $a + b$.   But $(m,x \in F(m))$ is precisely an object of the Grothendieck category $\inta F$, and the opfibration $U \maps \inta F \to \A$ sends this object to $m$.  Thus, $(e \to m, x \in F(m))$ can be seen as an object of the comma category $\A/U$.   Let $\Phi \maps \A/U \to \A$ be the functor sending any object $(e \to m,x \in F(m)) \in \A/U$ to $e$.  Then an $F$-decorated cospan from $a$ to $b$ amounts to an object $c \in \A/U$ such that $\Phi(c) = a + b$.    With this choice of $\Phi$, Shulman's result gives the double category of $F$-decorated cospans.

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$.  To obtain the desired double category we apply Shulman's \cref{lem:shulman} to the functor $\Phi \maps \A/U \to \A$, which we now describe in more detail.  In the comma category $\A/U$:
\begin{itemize}
\item an object is a pair $(a \xrightarrow{i} b, x)$ consisting of a morphism in $\A$ and an
object $x \in F(b)$;
\item a morphism from $(a \xrightarrow{i} b, x)$ to $(a' \xrightarrow{i'} b', x')$ is a triple $(f,g,h)$ where $f \maps a \to a'$ and $g \maps b \to b'$ are morphisms in $\A$ such that this square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
and $h \maps F(g)(x) \to x'$ is a morphism in $F(b')$.
\end{itemize}
As in any comma category we have a functor $\Phi \maps \A/U \to \A$; this is given on objects by $\Phi(a \xrightarrow{i} b, x)=a$ and on morphisms by $\Phi(f,g,h)=f$.

To apply \cref{lem:shulman} it suffices to prove that $\Phi$ is a Beck--Chevalley monoidal opfibration.   First, there is a monoidal structure on $\A/U$ coming from the monoidality of $U$.  On objects this is given by
\[  (a \xrightarrow{i} b, x) \otimes (a' \xrightarrow{i'} b', x') = (a+a' \xrightarrow{i+i'} b+b', 
\phi_{b,b'}(x,x') )\]
where $\phi$ is the laxator for $F$.  The monoidal unit is $(0_\A \xrightarrow{!} 0_\A, 0)$
where $0 \in F(0_\A)$ is the object given by $\phi_0 \maps \one \to F(0_\A)$.   With this monoidal structure it is easy to see that $\Phi$ is strict monoidal.

Next, one can check that $\Phi$ is a fibration.  Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a' \xrightarrow{i'} b', x') \in \A/U$ over $a'$, a cartesian lifting of $f$ to this object is given by 
\[ (f,1,1) \maps (a \xrightarrow{i' f} b' , x')\; \to \; (a' \xrightarrow{i'} b', x'). \]
Denoting the cartesian lifting of $f$ to the object $z \in \A/U$ by $\mathrm{Cart}(f,z)$, the tensor product of $\A/U$ strictly preserves such cartesian liftings in the sense that $\mathrm{Cart}(f_1,z_1) \otimes \mathrm{Cart}(f_2,z_2)=\mathrm{Cart}(f_1 \otimes f_2, z_1 \otimes z_2)$.

>>>>

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $(F,\phi,\phi_0) \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $s \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a \define{decoration morphism} $\tau \maps F(h)(s) \to s'$ in $F(s')$.
\end{itemize}
\end{thm}

@@

\[ \left(a\to m\leftarrow b,s\in F(m)\right), \qquad \left(b\to n\leftarrow c,t\in F(n)\right)\] 

@@

together with the decoration $t \odot s \in F(m +_b n)$ specified by this functor:
\[  \one\cong\one \times \one \xrightarrow{s \, \times \, t} F(n) \times F(m) \xrightarrow{\phi_{n,m}} F(m+n) \xrightarrow{F(\psi)} F(m +_{b}n). \]

@@

\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};

@@

\node (J) at (7,0.5) {$t \in F(n)$};
\node (K) at (7,-0.5) {$t' \in F(n')$};
\node (L) at (1,-1) {$\tau_\alpha \maps F(h_1)(s) \to s'$};
\node (M) at (5,-1) {$\tau_\beta \maps F(h_2)(t) \to t'$};

@@

\node (E) at (4.5,0.5) {$t \odot s \in F(m+_b n)$};
\node (F) at (4.5,-0.5) {$t' \odot s' \in F(m' +_{b'} n')$};

@@

together with the decoration morphism $\tau_{\beta \odot \alpha} \maps F(h_1 +_g h_2)(t \odot s) \to (t' \odot s')$ specified by this natural transformation:

@@

(D) edge node [above] {$s \times t$} (E)
(D) edge node [below] {$s' \times t'$} (E')

@@

where the middle isomorphism is \cref{eq:pseudonaturality} from the pseudonaturality of $\phi$ and the right-hand isomorphism comes from the pseudofunctoriality of $F$.

\begin{thm}\label{thm:decorated_cospans_2}
Let $(\A,+)$ be a cocartesian monoidal category and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ be a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a \to m \leftarrow b, s \in F(m))$ and $(c \to n \leftarrow d, t \in F(n))$ is
\[
\begin{tikzcd}
& m+n & \\
a+c\ar[ur,"i_1+i_2"] && b+d,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad s \otimes t:=  \phi_{m,n}(s,t) \in F(m + n) 
\]
\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(s) \to s' \textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(t) \to t'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m,n}(s,t))\to \phi_{m',n'}(s',t')\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{h,h'}}{\cong}}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"s\,\times \, t"]\ar[dr,"s' \,\times \, t' "']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

We prove both these theorems using the work of Shulman \cite{Shulman2008}, who gives a general way to construct a double category from a `Beck--Chevalley monoidal bifibration'---a concept recalled in \cref{subsec:fibrations}.

@@

Note that when $\C$ is the arrow category of $\A$ and $\Phi$ maps any arrow in $\C$ to its
domain, an object of $\Fr(\Phi)$ simply amounts to a cospan in $\A$.  In this case $\Fr(\Phi)$ is the double category of cospans in $\A$.    We shall use variations on this idea to construct double categories of decorated and structured cospans, and to prove that under certain conditions these double categories are equivalent.    We begin with decorated cospans, proving Theorem \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying Shulman's result to a particular monoidal bifibration built from a lax monoidal pseudofunctor $F \maps (A,+) \to (\Cat,\times)$.

@@

In more detail, if $(\phi,\phi_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{subsec:bicats}, the induced monoidal structure on the Grothendieck category $\X=\inta F$ (\cref{def:GrothCat}) is given by
\begin{equation}\label{eq:explicitstructure2}
\Big(a,s\in F(a)\Big)\ot_{\X}\Big(b,t\in F(b)\Big)=\Big(a+b,\phi_{a,b}(s,t)\in F(a+b)\Big), \qquad I_{\X}=\Big(0_A,\phi_0\Big)
\end{equation}
If $F$ is a symmetric lax monoidal pseudofunctor, the induced monoidal structure in $\inta F$ is symmetric via
$$\left(\beta_{a,b},(u_{a,b})_{s,t}\right)\maps \left(a+b,\phi_{a,b}(s,t)\right)\simrightarrow\left(b+a,\phi_{b,a}(t,s)\right)$$ where $\beta$ is the canonical symmetry for $\A$ and $u$ is the natural isomorphism of \cref{eq:braidedpseudofun}.

Moreover, each fiber $\X_a=F(a)$ obtains a monoidal structure via 
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\phi_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a)
\end{equation}
where $\nabla$ is the fold map, which is symmetric when $F$ is, again via the components of $u_{a,a}$. Also, each reindexing functor $f_!=F(f)$ obtains a strong monoidal structure with
\begin{equation}\label{eq:strongreindexing}
\begin{tikzcd}[row sep=.35in]
 F(a)\times F(a)\ar[r,"\phi_{a,a}"]\ar[d,"F(f)\times F(f)"']\ar[dr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{f,f}}{\cong}}"] & 
 F(a+a)\ar[r,"F(\nabla)"]\ar[d,"F(f+f)"description]\ar[dr,phantom,"\cong"] & F(a)\ar[d,"Ff"] \\
 F(b)\times F(b)\ar[r,"\phi_{b,b}"'] & F(b+b)\ar[r,"F(\nabla)"'] & F(b)
\end{tikzcd}
\end{equation}
\end{proof}

Now we are ready to use Shulman's result to prove Theorems \ref{thm:decorated_cospans_1} and  \ref{thm:decorated_cospans_2}.  An $F$-decorated cospan from $a$ to $b$ 
is a cospan $a \to m \leftarrow b$ in $\A$ together with an object $s \in F(m)$.   This can be seen as a pair $(e \to m, s \in F(m))$ such that $e$ equals $a + b$.   But $(m,s \in F(m))$ is precisely an object of the Grothendieck category $\inta F$, and the opfibration $U \maps \inta F \to \A$ sends this object to $m$.  Thus, $(e \to m, s \in F(m))$ can be seen as an object of the comma category $\A/U$.   Let $\Phi \maps \A/U \to \A$ be the functor with
\[    \Phi(e \to m,s \in F(m)) = e .\]
Then an $F$-decorated cospan from $a$ to $b$ amounts to an object $c \in \A/U$ such that $\Phi(c) = a + b$.    With this choice of $\Phi$, Shulman's result gives the double category of $F$-decorated cospans.

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$.  To obtain the desired double category we apply Shulman's \cref{lem:shulman} to the functor $\Phi \maps \A/U \to \A$, which we now describe in more detail.  In the comma category $\A/U$:
\begin{itemize}
\item an object is a pair $(a \xrightarrow{i} b, s)$ consisting of a morphism in $\A$ and an
object $s \in F(b)$;
\item a morphism from $(a \xrightarrow{i} b, s)$ to $(a' \xrightarrow{i'} b', s')$ is a triple $(f,g,h)$ where $f \maps a \to a'$ and $g \maps b \to b'$ are morphisms in $\A$ such that this square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
and $h \maps F(g)(s) \to s'$ is a morphism in $F(b')$.
\end{itemize}
As in any comma category we have a functor $\Phi \maps \A/U \to \A$; this is given on objects by $\Phi(a \xrightarrow{i} b, x)=a$ and on morphisms by $\Phi(f,g,h)=f$.

To apply \cref{lem:shulman} it suffices to prove that $\Phi$ is a Beck--Chevalley monoidal opfibration.   First, there is a monoidal structure on $\A/U$ coming from the monoidality of $U$.  On objects this is given by
\[  (a \xrightarrow{i} b, s) \otimes (a' \xrightarrow{i'} b', s') = (a+a' \xrightarrow{i+i'} b+b', 
\phi_{b,b'}(s,s') )\]
where $\phi$ is the laxator for $F$.  The monoidal unit is $(0_\A \xrightarrow{!} 0_\A, 0)$
where $0 \in F(0_\A)$ is the object given by $\phi_0 \maps \one \to F(0_\A)$.   With this monoidal structure it is easy to see that $\Phi$ is strict monoidal.

Next, one can check that $\Phi$ is a fibration.  Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a' \xrightarrow{i'} b', s') \in \A/U$ over $a'$, a cartesian lifting of $f$ to this object is given by 
\[ (f,1,1) \maps (a \xrightarrow{i' f} b' , s')\; \to \; (a' \xrightarrow{i'} b', s'). \]
Denoting the cartesian lifting of $f$ to an object $z \in \A/U$ by $\mathrm{Cart}(f,z)$, the tensor product of $\A/U$ strictly preserves such cartesian liftings in the sense that $\mathrm{Cart}(f_1,z_1) \otimes \mathrm{Cart}(f_2,z_2)=\mathrm{Cart}(f_1 \otimes f_2, z_1 \otimes z_2)$.
